---
title: "Creating reproducible packages when data are confidential: Bonus slides"
author: 
  - "Lars Vilhuber"
date: "`r Sys.Date()`"
output: 
  ioslides_presentation:
    incremental: false
    self-included: true
    widescreen: true
---


# Use cross-platform programming practices

## Cross-platform programming practices 1

**Use programming-language specific code as much as possible**

Avoid
```{r, eval=FALSE}
system("unzip C:\data\myfile.zip")
```
or
```{stata, eval=FALSE}
shell unzip "C:\data\myfile.zip"
```


## Cross-platform programming practices 1

Most languages have appropriate code:

R:

```{r, eval=FALSE}
unzip(zipfile, files = NULL, list = FALSE, overwrite = TRUE,
      junkpaths = FALSE, exdir = ".", unzip = "internal",
      setTimes = FALSE)
```

Stata:

```{stata, eval=FALSE}
unzipfile "zipfile.zip" [, replace]
```


## Cross-platform programming practices 2

Use neutral pathnames (mostly forward slashes)

::: {.columns-2}


:::: {.column}

**R**: Use functions to combine paths (and/or use forward slashes), packages to make code more portable.

<div class="red2">
```
basepath <- rprojroot::find_root(rprojroot::has_file("README.md"))
data <- read.dta(file.path(basepath,"path","data.dta"))
```
</div>
::::

:::: {.column}

**Stata**: *always* use forward slashes, even on Windows

<div class="blue2">

```
global data "/my/computer"
use "$data/path/data.dta"
```
</div>

::::

:::
